\documentclass[letterpaper, 11pt]{extarticle}\usepackage[]{graphicx}\usepackage[]{xcolor}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlsng}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hldef}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\input{preamble}
\input{format}
\input{commands}
\usepackage{newpxmath}
\usepackage[margin=0.95in]{geometry}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{xcolor}
\definecolor{softpurple}{HTML}{9C88B9}
\definecolor{softblue}{HTML}{B8BDDE}
\usepackage{cancel}
\usepackage{csquotes}
\usepackage{enumitem}

\definecolor{darkblue}{RGB}{13, 39, 61}  % Yankees Blue
\definecolor{mediumblue}{RGB}{62, 105, 133}  % Teal Blue
\definecolor{lightblue}{RGB}{138, 167, 188}  % Pewter Blue
\definecolor{pastelblue}{RGB}{166, 190, 209}  % Pastel Blue
\definecolor{columbiablue}{RGB}{205, 215, 223}  % Columbia Blue
\definecolor{graytext}{RGB}{90, 90, 90}  % Neutral grey for comments
\usepackage{listings}
\lstset{
  language=R,
  basicstyle=\ttfamily\small,
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1}


\lstset{ 
  language=R,
  basicstyle=\ttfamily\small\color{darkblue},
  numbers=left,
  numberstyle=\tiny\color{gray},
  stepnumber=1,
  numbersep=5pt,
  backgroundcolor=\color{white},
  showspaces=false,
  showstringspaces=false,
  breaklines=true,
  breakatwhitespace=true,
  captionpos=b,
  escapeinside={\%*}{*)},
  commentstyle=\color{graytext},
  keywordstyle=\color{softpurple},
  stringstyle=\color{pastelblue}
}

\setcitestyle{square}

\usepackage{color}
\usepackage{framed}
\newcommand{\sam}[1]{
  {\color{blue} [SAM: {#1}]}
 }
 \newcommand{\asl}[1]{
  {\color{blue} [ASL: {#1}]}
 }

 \definecolor{nyscolor}{HTML}{a63a47}

\newcommand{\nys}[1]{%
  {\color{nyscolor} [NYS: {#1}]}%
}

 \newcommand{\nys}[1]{
  {\color{blue} [NYS: {#1}]}
 }

% Custom colors
\definecolor{shadecolor}{RGB}{255,255,255}  % white background
\definecolor{Rcode}{RGB}{136, 18, 128}      % keyword purple (similar to print)
\definecolor{Rout}{RGB}{90, 90, 90}         % output gray
\definecolor{Rcomment}{RGB}{110,110,110}    % comment gray

% Redefine how R chunks are styled
\usepackage{etoolbox}
\makeatletter
\patchcmd{\@Verbatim}{\fbox}{\color{black}}{}{}
\makeatother

% Required if using knitr
\usepackage{alltt}
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl,formatcom=\color{Rcode},commandchars=\\\{\}}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{formatcom=\color{Rout}}
\DefineVerbatimEnvironment{Scode}{Verbatim}{formatcom=\color{Rcode}}

\geometry{letterpaper}                  
\hypersetup{colorlinks=false, allcolors=blue}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\begin{Large}
    \textsf{\textbf{Empirical Bayes: Step Count Problem}}   \par
    Statistical Reinforcement Learning Lab
\end{Large}

\vspace{2ex}

\textsf{\textbf{Problem Statement}}

In our study, we have a population of 30 unique individuals and are interested in investigating the number of steps each person takes in a 30-minute interval. We designate the true mean steps for each of 30 individuals as $\mu_1, \mu_2, \dots, \mu_i$ for $i=1, 2,\dots, 30$  ($n=30$). 

During the study, we use a device to measure the number of steps each individual takes during a 30-minute interval. We use the device to perform this measurement once for each individual. 

Now, we have two data points for each individual:\begin{enumerate}
\itemsep0em 
    \item The device's step count for that individual.
    \item A measure of uncertainty provided by the device, which describes how sure the device is that it counted the number of steps correctly.
\end{enumerate}
We designate the observed step counts as $x_1, x_2, \dots, x_i$ and the measures of uncertainty as $\sigma_1, \sigma_2, \dots, \sigma_i$. Thus, our data for each of the 30 individuals consists of one pair of values, $(x_i, \sigma_i)$. 

The true means and hyperparameters are unknown to us. The observed step counts and measures of uncertainty are our data and thus are known to us. How can we estimate $\mu_0$ and $\tau^2$ using this data?
\\

\textsf{\textbf{Empirical Bayes Solution}}

We will approach this problem using Raudenbush \& Bryk's empirical Bayes model for meta-analyses \cite{doi:10.3102/10769986010002075}. 

First, we assume that each person's true mean steps is a draw from the same underlying normal distribution with some non-zero variance $\tau^2$: $$\mu_i \sim \mathcal{N}(\mu_0, \tau^2).$$ 

We also assume each person's step count on a given day is a draw from a normal distribution centered at that person's mean, $\mu_i$, with non-zero variance $\sigma_i^2$. (We assume that the device's uncertainty measurement, $\sigma_i^2$, accurately captures the true within-person variance.)

Then, for this problem, $\tau^2$ designates population variance while $\sigma_i^2$ designates within-person variance. 

By equation 8 (p. 78) and equation 18 (p. 80) \cite{doi:10.3102/10769986010002075}, we can model each individual's observation is a draw from a normal distribution centered at that person's mean $\mu_i$. The variance of this observation is the sum of within-person and population variance. Therefore, the distribution of the observed data $\textbf{x}$ is:
$$x_i \sim \mathcal{N}(\mu_0, \sigma_i^2+\tau^2), \quad x_i \mid \mu_i \sim \mathcal{N}(\mu_i, \sigma_i^2).$$

Let us note that our model parallels Raudenbush \& Bryk's model as follows \cite{doi:10.3102/10769986010002075}: $d_i = x_i$, $W_i' = 1$, $\gamma = \mu_0$, $v_i = \sigma_i^2$. Recall that in our problem, $\sigma_i^2$ is known to us as it is provided by the device, but $\tau^2$ and $\mu_0$ are both unknown. Thus, we are in the scenario with unknown variance and will reference the \textit{Variance estimation} section (pp. 80-82). 

We can derive the likelihood and log-likelihood functions of $\tau^2$ and $\mu_0$ from the normal PDF as follows:
\begin{align}
L(\mu_0, \tau^2)
&= \prod_{i=1}^{n}
  \mathcal{N}\!\left(x_i \mid \mu_0,\, \sigma_i^2 + \tau^2\right)\\
L(\mu_0, \tau^2)
&= \prod_{i=1}^{n}
  \frac{1}{\sqrt{2\pi(\sigma_i^2 + \tau^2)}}
  \exp\!\left[
    -\frac{(x_i - \mu_0)^2}{2(\sigma_i^2 + \tau^2)}
  \right]\\
  \ell(\mu_0, \tau^2)
&= -\frac{1}{2} \sum_{i=1}^{n}
  \left\{
    \log(\sigma_i^2 + \tau^2)
    + \frac{(x_i - \mu_0)^2}{\sigma_i^2 + \tau^2}
  \right\}
\end{align}
We can rewrite the conditional log-likelihood as follows:
\begin{align}
Q(\mu_0;\tau^2)=\sum_{i=1}^{n}w_i(\tau^2)\,(x_i-\mu_0)^2,
\quad w_i(\tau^2)=\frac{1}{\sigma_i^2+\tau^2}
\end{align}
Next, we differentiate w.r.t. $\mu_0$ to determine the MLE estimator for $\mu_0$ conditional on $\tau^2$. Here, we are considering $\tau^2$ as a constant and hence the $\log(\sigma_i^2 + \tau^2)$ term is dropped.
\begin{align}
    \frac{\partial Q}{\partial \mu_0}
&=-2\sum_{i=1}^{n}w_i(\tau^2)\,(x_i-\mu_0)=0\\
\mu_0^*(\tau^2)&=\frac{\sum_{i=1}^{n} w_i(\tau^2)\,x_i}{\sum_{i=1}^{n} w_i(\tau^2)}
\end{align}
We have effectively written the conditional MLE of $\mu_0$ in terms of $\tau^2$, so we only have one unknown variable remaining ($\tau^2$). We will replace $\mu_0$ with our conditional MLE in the original log-likelihood equation. Now, we have:
\begin{align}
    \ell_p(\tau^2)
= -\frac12\left\{
\sum_{i=1}^{n}\log(\sigma_i^2+\tau^2)
	+	Q_1(\tau^2)
\right\}, \quad Q_1(\tau^2)=\sum_{i=1}^{n} w_i(\tau^2)\,\bigl(x_i-\mu_0^*(\tau^2)\bigr)^2
\end{align} 

We define the MLE of $\tau^2$ as:
\begin{align}
    \ \hat{\tau}^2=\arg\max_{\tau^2\ge 0}\ \ell_p(\tau^2)\
\end{align}
Now, we can substitute this estimate back into our conditional MLE estimator for $\mu_0$ to obtain an estimate for $\mu_0$:
\begin{align}
    \ \hat{\mu}_0=\mu_0^*(\hat{\tau}^2)
=\frac{\sum_{i=1}^{n} (\sigma_i^2+\hat{\tau}^2)^{-1} x_i}
{\sum_{i=1}^{n} (\sigma_i^2+\hat{\tau}^2)^{-1}}\
\end{align}
Now, we can write the posterior likelihood function for $\mu_i$ as follows:
\begin{align}
L(\mu_i; \mathbf{x})
&=\prod_{i=1}^{n}
\frac{1}{\sqrt{2\pi\sigma_i^2}}
\exp\!\left\{-\frac{(x_i-\mu_i)^2}{2\sigma_i^2}\right\}
\cdot
\frac{1}{\sqrt{2\pi\tau^2}}
\exp\!\left\{-\frac{(\mu_i-\mu_0)^2}{2\tau^2}\right\}\\
\ell(\mu_i; \mathbf{x})
&= -\frac{1}{2}\sum_{i=1}^{n}\Bigg[
\log(2\pi\sigma_i^2)+\frac{(x_i-\mu_i)^2}{\sigma_i^2}
+\log(2\pi\tau^2)+\frac{(\mu_i-\mu_0)^2}{\tau^2}
\Bigg]
\end{align}
First, we take the partial derivative w.r.t. $\mu_i$ to derive the posterior mean, substituting in the MLE estimators we derived above.
\begin{align}
    &\frac{\partial \ell(\mu_i; \mathbf{x})}{\partial \mu_i}
= \frac{x_i-\mu_i}{\sigma_i^2}+\frac{\mu_0-\mu_i}{\tau^2}\\
&\Rightarrow\ \mu_i\left(\frac{1}{\sigma_i^2}+\frac{1}{\hat{\tau}^2}\right)
=\frac{x_i}{\sigma_i^2}+\frac{\hat{\mu_0}}{\hat{\tau}^2}
\\&\Rightarrow
\ \mu_i
=\lambda_i x_i+(1-\lambda_i)\hat{\mu_0}, \quad 
\lambda_i=\frac{\hat{\tau}^2}{\sigma_i^2+\hat{\tau}^2}
\end{align}
Second, we derive the MLE of the posterior variance by Fisher Information:
\begin{align}
\frac{\partial^2 \ell}{\partial \mu_i^2} 
&= -\left( \frac{1}{\sigma_i^2} + \frac{1}{\tau^2} \right) \\[8pt]
\Rightarrow\quad 
\operatorname{Var}(\mu_i \mid x_i) 
&= \left( -\frac{\partial^2 \ell}{\partial \mu_i^2} \right)^{-1} 
= \left( \frac{1}{\sigma_i^2} + \frac{1}{\tau^2} \right)^{-1}
\end{align}
To conclude, we have:
\begin{align}
    \textrm{Prior: }&\mu_i \sim \mathcal{N}(\mu_0, \tau^2)\\
    \textrm{Posterior: }& \mu_i \mid \mathbf{x} \sim \mathcal{N}\left(
\lambda_i x_i+(1-\lambda_i)\hat{\mu_0},\;
\Big(\frac{1}{\sigma_i^2}+\frac{1}{\hat{\tau}^2}\Big)^{-1}
\right), \quad \lambda_i=\frac{\sigma_i^2}{\sigma_i^2+\hat{\tau}^2}.
\end{align}
\newpage
\bibliographystyle{plain}
\bibliography{bib}

\end{document}

%px1 x
%pxp matrix of uncertainity for each person

%everyone's x is 1x1 MAB code & consult with annie 
